---
title: "Comparing actxps and ExperienceAnalysis.jl"
output:
  html_document:
    df_print: paged
---

We do a comparison of exposures created by actxps (an R package) and ExperienceAnalysis.jl (a Julia package).

```{r, message=FALSE}
library(actxps)
library(readr)
library(magrittr)
library(dplyr)
library(lubridate)
```

## Different row counts

```{r, message=FALSE}
census_dat <- read_csv("census_dat.csv")
r_df <- expose_py(
  census_dat,
  start_date = "2006-6-15",
  end_date = "2020-02-29",
  target_status = "Surrender"
) %>% select(pol_num, pol_date_yr, term_date, exposure, status)
jl_df <- read_csv("df_jl.csv") # from create_csv.jl
```

ExperienceAnalysis.jl creates 1887 more rows of exposures. We want to understand why.

```{r}
print(paste("row count Julia", nrow(jl_df)))
print(paste("row count R", nrow(r_df)))
print(paste("difference", nrow(jl_df)-nrow(r_df)))
```

## R left join Julia

We can use left joins to find rows from R that have no match in Julia.

```{r}
r_julia <- r_df %>% left_join(jl_df, c("pol_num", "pol_date_yr" = "from"))
```

### ExperienceAnalysis.jl generates duplicate exposures {#jl-dup}

The warning above lets us know that there are multiple matches, indicating duplicate (`policy_num`, `from`) combinations from ExperienceAnalysis.jl. We see that ExperienceAnalysis.jl generates three rows with no `exposure_fraction`.

```{r}
jl_df %>%
  group_by(pol_num, from) %>%
  mutate(pol_from_count=n()) %>%
  filter(pol_from_count > 1)
```

### Inspect rows with no match

Rows from actxps that have no match in ExperienceAnalysis.jl follow these patterns:

* If `term_date` is defined, `term_date == pol_date_yr`
* If `term_date` is not defined
  * `pol_date_yr` falls on a leap day (xxxx-02-29) or
  * `pol_date_yr` falls on 2020-03-01

```{r}
r_julia %>% filter(is.na(exposure_fraction))
```

### `term_date` defined, `term_date == pol_date_yr` {#term-is-anniv}

ExperienceAnalysis.jl appears to treat date intervals with a non-inclusive right boundary, [issue_date, termination_date). actxps appear to have an inclusive right boundary.

```{r}
r_julia %>% filter(pol_num %in% c(640, 1523))
```

According to section 4.3 of the Society of Actuaries (SOA) [experience study document](https://www.soa.org/49378a/globalassets/assets/files/research/experience-study-calculations.pdf), both of these approaches are wrong some of the time.

> For a lapse on a policy anniversary, using 11:59 pm on the day
before the anniversary assures that the lapse is allocated to the proper policy year. The date assumption may need
to be adjusted for certain events under study. For example, a death on the policy anniversary would be incorrectly
assigned to the prior policy year by using 11:59 on the day before. Deaths should therefore be assumed to occur at
11:59 pm on the date of death, not the prior day.

ExperienceAnalysis.jl is not correct on `pol_num` 640 because it does not create an exposure interval containing the day 2014-11-02. actxps is not correct on `pol_num` 1523 because it assigns the lapse to the day 2019-09-30 instead of 2019-09-29.

### `term_date` not defined, `pol_date_yr` falls on a leap day (`xxxx-02-29`) {#leap}

actxps does not create exposures properly for policies issued on leap day.

```{r}
r_df %>% filter(pol_num == 10465)
```
ExperienceAnalysis.jl appears to not assign some dates to the correct interval. The fifth row should start on 2012-02-29.

```{r}
jl_df %>% filter(pol_num == 10465)
```

### `term_date` not defined, `pol_date_yr` falls on 2020-03-01 {#right-trunc}

The end date of the study is 2020-02-29, so this should not happen. I am unsure if this is related to having an end date that falls on a leap year.

```{r}
r_julia %>% filter(pol_num %in% c(2830,2877,1397,4621))
```

## Julia left join R

We do the same inspection of rows with no match.

```{r}
julia_r <- jl_df %>% left_join(r_df, c("pol_num", "from" = "pol_date_yr"))
julia_r %>% filter(is.na(exposure))
```

The rows in Julia that are not in R all have `from` as 2006-06-15 or xxxx-02-28.

```{r}
julia_r %>%
  filter(is.na(exposure)) %>%
  group_by(from) %>%
  summarise(count=n())
```

Rows of the form xxxx-02-28 are explained in the [previous section](#leap) on leap days.

### `from` is 2006-06-15 {#left-trunc}

Policy 4120 was issued on date 2005-05-27. The start date of the study truncates the interval [2006-05-27, 2007-05-27) to [2006-06-15,	2007-05-27). This appears to work as expected in ExperienceAnalysis.jl.

```{r}
jl_df %>% filter(pol_num == 4120)
```

actxps appears to not create partial exposure intervals that begin at the start date of the study.

```{r}
r_df %>% filter(pol_num == 4120)
```

## Review

The following issues will be made on GitHub:

* ExperienceAnalysis.jl
  * Creation of empty exposures, [link](#jl-dup)
  * Handling of deaths that occur on policy anniversaries, [link](#term-is-anniv)
  * Leap day bug is cause by the following: (((Leap Day + Year) + Year) + Year) + Year != Leap Day + 4 Years. Change way that starting interval dates are created. [link](#leap)
* actxps
  * Handling of surrenders that occur on policy anniversaries, [link](#term-is-anniv)
  * Create partial exposure intervals that begin at the start date of the study, [link](#left-trunc)
  * Exposure intervals created past study end, [link](#right-trunc)
  * Policies issued on leap year do not create exposures on non-leap years, [link](#leap)
  
